---
title: "Imperfect Detection"
author: "Michelle DePrenger-Levin"
date: "2024-06-03"
output: html_document
---

Dormancy and Detection

Load libraries   
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMark)
library(popbio)
library(RCurl)
library(devtools)
library(HDInterval)
library(Rage)
```


Matrix population model for Astragalus microcymbus    
```{r}

## Get 2 stage PVA to make MPMs
source_url("https://raw.githubusercontent.com/DenverBotanicGardens/PVA/master/StagePVA2stages_dormancyoptional.R")

## Load the stage-fate dataset
load("data/asmi.Rdata")

## StagePVA2stages_dormancyoptional.R (dormancy isn't optional yet)
MPM.plots <- StagePVA(asmi.all2,TF = FALSE)
MPM.plotsSplit <- StagePVA(asmi.all2,TF = TRUE)

SitePlot <- asmi.all2 %>%
  distinct(plot,site) %>%
  mutate(plot = as.factor(plot))%>%
  mutate(site = as.factor(site))
```


Compare to Pradel model   
```{r}
load("data/asmi.raw")

Pradel_asmi <- asmi.raw %>%
  filter(!(AsMi_tag_id %in% c(1177,1171,2746,3903))) %>%
  filter(!is.na(year)) %>%  
  filter(AsMi_site_id %in% c(5,15,19,26)) %>% ## exclude Cebolla Creek that was added 2014
  dplyr::select(c(AsMi_tag_id:status,AsMi_plot_id:fence)) %>%
  group_by(AsMi_tag_id) %>%
  filter(any(length > 0)) %>%
  ungroup() %>%
  mutate(Obs = case_when(length > 0 ~ 1, 
                         length == 0 ~ 0,
                         is.na(length) ~ 0)) %>%
  arrange(year) %>%
  mutate(Obs = as.character(Obs)) %>%
  tidyr::pivot_wider(names_from = year, values_from = Obs, values_fill = "0", names_prefix = "Year", 
              id_cols = c(AsMi_tag_id,AsMi_site_id,AsMi_plot_id)) %>%
  tidyr::unite(ch, Year1995:Year2022, sep ="") %>%
  mutate(Tag = as.factor(AsMi_tag_id),
         Site = as.factor(AsMi_site_id),
         Plot = as.factor(AsMi_plot_id)) %>%
  filter(grepl("1",ch))

dmPlot <- model.matrix(~ -1 + Plot, Pradel_asmi)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]
plotPradeldf <- Pradel_asmi %>%
  distinct(Plot)


Pradel_asmi %>%
  dplyr::select(ch) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot799 = paste(Plot799, ";", sep = "")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", sep = " ", 
              col.names = FALSE, row.names = FALSE)


asmiPradel.inp <- convert.inp("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Pradel_asmi_mark.inp", 
                        group.df = plotPradeldf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiproc <- process.data(asmiPradel.inp, model = "Pradrec", groups = "Plot", begin.time = 1995)
asmiddl <- make.design.data(asmiproc)

## Fix p to zero for plots not surveyed 
### filter(AsMi_plot_id != 598) %>% ## exclude the one plot added in 2004
### filter(!(AsMi_plot_id %in% c(300,238))) %>% ## Since two plots, 300, 238 were added in 1996
pindex <- asmiddl$p
p598indexApp3 <- as.numeric(row.names(asmiddl$p[(asmiddl$p$time %in% c("1995","1996","1997","1998","1999",
                                                                      "2000","2001","2002","2003") &
                                                   asmiddl$p$group == "598"),]))
p598idx <- pindex %>%
  mutate(time = as.numeric(as.character(time)))%>%
  filter(time < 2004 & group == "598") %>%
  mutate(idx = as.numeric(row.names(.)))
p300_238idx <- pindex %>%
  filter(time == 1995 & group %in% c("300","238")) %>%
  mutate(idx = as.numeric(row.names(.)))


run.asmi <- function(){
  Phi.dot = list(formula =  ~ 1)
  Phi.time = list(formula =  ~ time)
  Phi.Plot = list(formula =  ~ Plot)  ## so each plot is offset
  Phi.timePlot = list(formula =  ~ time + Plot)
  
  p.dot = list(formula =  ~ 1)
  # p.time = list(formula = ~ time)
  # p.Plot = list(formula = ~ -1 + Plot)
  # p.timePlot = list(formula = ~ time + Plot)
  
  f.dot = list(formula =  ~ 1)
  f.time = list(formula = ~ time)
  f.Plot = list(formula = ~ Plot)
  f.timePlot = list(formula = ~ time + Plot)
  
  asmi.model.list <- create.model.list("Pradrec")
  
  asmi.results <- mark.wrapper(asmi.model.list, data = asmiproc, ddl = asmiddl)
  
  return(asmi.results)
}

asmi.models <- run.asmi()
```



Compare multi-state mark-recapture to MPM estimates of survival, transitions    
```{r}

## dataCJS
load("data/dataCJS3yrs.Rdata")

## Can only be one strata within a primary session - need to assign V if never reproductive, R if eventually reproductive  
##### Could condense to spring and fall, then have V early and some R late
## Repeated for the three years
obsmodel <- dataCJS %>%
  filter(Site %in% c(5,15,19,26)) %>%
  group_by(Year, Plot, Site, Tag) %>%
  ## Make the multistate options
  dplyr::mutate(MSObs = case_when(Ln == 0 ~ "0",
                                  (Ln > 0 & Fl == 0) ~ "V",
                                  any((Ln > 0 & Fl == 1)) ~ "R")) %>%
  # dplyr::mutate(MSObs = case_when(any(MSObs == "R") ~ replace(MSObs, MSObs == "V", "R"),
  #                                 TRUE ~ MSObs)) %>%
  ungroup()%>%
  distinct(Day,Year,Plot,Site,Tag,encounter, .keep_all = TRUE) %>%
  # dplyr::select(c(Site:Tag, encounter, Obs, Year)) %>%
  dplyr::select(c(Site:Tag, encounter, MSObs, Year)) %>%
  ## Fill in all tags within a Plot and Site across all years
  tidyr::complete(nesting(Site,Plot,Tag),Year, encounter, fill = list(MSObs = "0")) %>%
  arrange(encounter) %>%
  pivot_wider(names_from = encounter, values_from = MSObs, values_fill = "0", names_prefix = "Occ") 

obsmodel %>%
  filter(if_any(.cols = everything(), ~ grepl("V", .))) %>%
  print(n=100)

dataCJS <- dataCJS %>% filter(Site %in% c(5,15,19,26))
plotssurveyed <- table(dataCJS$Plot, dataCJS$encounter, dataCJS$Year)  

## In the design data lists after processing data, set these to zero detection (p)


asmiRD <- obsmodel %>%
  ## unite within each year
  tidyr::unite(ch, Occ1:Occ10, sep="") %>%
  ## Change all to R if any within a year
  dplyr::mutate(ch = if_else(grepl("R",ch), gsub("V","R",ch), ch)) %>%
  dplyr::group_by(Site,Plot,Tag) %>%
  dplyr::mutate(Plot = as.factor(Plot)) %>%
  dplyr::mutate(Site = as.factor(Site)) %>%
  ## paste across years
  dplyr::summarise(ch = paste0(ch, collapse = "")) %>%
  ungroup() %>%
  filter((grepl("V", ch) | grepl("R",ch)))

asmiRD %>%
  filter(grepl("R",ch))

dmPlot <- model.matrix(~ -1 + Plot, asmiRD)
dmSite <- model.matrix(~ -1 + Site, asmiRD)
dimnames(dmPlot)[[2]][length(dimnames(dmPlot)[[2]])]
dimnames(dmSite)[[2]][length(dimnames(dmSite)[[2]])]

## By Plot
asmiRD %>%
  ungroup() %>%
  dplyr::select(c(ch)) %>%
  bind_cols(dmPlot) %>%
  mutate(Plot614 = paste(Plot614, ";", sep ="")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Robust3yearsPlot.inp", 
              sep = " ", col.names = FALSE, row.names = FALSE)

## By Site
asmiRD %>%
  ungroup() %>%
  dplyr::select(c(ch)) %>%
  bind_cols(dmSite) %>%
  mutate(Site26 = paste(Site26, ";", sep ="")) %>%
  write.table(file = "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/GitPVA/datasets/Robust3yearsSite.inp", 
              sep = " ", col.names = FALSE, row.names = FALSE)


plotdf <- asmiRD %>%
  ungroup()%>%
  dplyr::distinct(Plot)

sitedf <- asmiRD %>%
  ungroup() %>%
  dplyr::distinct(Site)
```

Mark-recapture by plot  
Mark-recpature by site
```{r}

## Convert for MARK  
asmiRDPlot.inp <- convert.inp("Robust3yearsPlot.inp", 
                        group.df = plotdf,
                        covariates = NULL,
                        use.comments = FALSE)

asmiRDSite.inp <- convert.inp("Robust3yearsSite.inp", 
                        group.df = sitedf,
                        covariates = NULL,
                        use.comments = FALSE)
```



# Huggins Closed Robust Design Multi-state with State Probabilities  
"CRDMSOHug"


We used a Robust design with a Huggins’ estimator in Program MARK (accessed through RMark ) to estimate survival and a Pradel Robust Model with survival and Lambda with Huggins’ closed capture estimator in Program MARK to estimate the population growth rate (Pollock, 1982; White et al., 2001). 


S(s): survival rate between primary occasions for state s 
Psi(r,s): transition between states between primary periods from r to state s
Omega: probability of being in each of the states
p(s,t): apparent detection
c(s,t): true detection   
```{r}

time.intervals <- c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0) 

RDproc <- process.data(asmiRDPlot.inp, model = "CRDMSOHug", begin.time = 2013, groups = "Plot",
                       time.intervals = time.intervals)
RDddl <- make.design.data(RDproc) #, parameters = list(Psi = list(subtract.stratum = c("R","R"))))

RDddl$Psi$stratum <- factor(RDddl$Psi$stratum, levels = c("R","V"))
RDddl$S$stratum <- factor(RDddl$S$stratum, levels = c("R","V"))
RDddl$Omega$stratum <- factor(RDddl$Omega$stratum, levels = c("R","V"))
RDddl$p$stratum <- factor(RDddl$p$stratum, levels = c("R","V"))
RDddl$c$stratum <- factor(RDddl$c$stratum, levels = c("R","V"))

table(RDddl$Psi[,c("stratum","tostratum")]) ## R to R and V to V computed by subtraction
## Now add zero for detection p, at all times when not visited; session = Year, time = secondary occasion 1:10, Plot = Plot
pindex <- RDddl$p
## Missed entire survey period
p2013.indices <- pindex %>%
  filter(session == 2013 & time %in% c(1,2,4,5,7,9))
## Missed these plots [Plot, Survey, Year]
p2013.ind3 <- pindex %>%
  filter(session == 2013 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,1])[which(plotssurveyed[,3,1] == 0)]))
p2013.ind6 <- pindex %>%
  filter(session == 2013 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,1])[which(plotssurveyed[,6,1] == 0)])) 
p2013.ind8 <- pindex %>%
  filter(session == 2013 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,1])[which(plotssurveyed[,8,1] == 0)])) 
p2013.ind10 <- pindex %>%
  filter(session == 2013 & time == 10 & Plot %in% as.numeric(names(plotssurveyed[,10,1])[which(plotssurveyed[,10,1] == 0)])) 
p2013.inds <- c(p2013.indices$par.index, p2013.ind3$par.index, p2013.ind6$par.index, p2013.ind8$par.index, p2013.ind10$par.index)

## Missed in 2014
p2014.indices <- pindex %>%
  filter(session == 2014 & time == 4)
p2014.ind1 <- pindex %>%
  filter(session == 2014 & time == 1 & Plot %in% as.numeric(names(plotssurveyed[,1,2])[which(plotssurveyed[,1,2] == 0)])) 
p2014.ind2 <- pindex %>%
  filter(session == 2014 & time == 2 & Plot %in% as.numeric(names(plotssurveyed[,2,2])[which(plotssurveyed[,2,2] == 0)])) 
p2014.ind3 <- pindex %>%
  filter(session == 2014 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,2])[which(plotssurveyed[,3,2] == 0)])) 
p2014.ind5 <- pindex %>%
  filter(session == 2014 & time == 5 & Plot %in% as.numeric(names(plotssurveyed[,5,2])[which(plotssurveyed[,5,2] == 0)]))
p2014.ind6 <- pindex %>%
  filter(session == 2014 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,2])[which(plotssurveyed[,6,2] == 0)])) 
p2014.ind7 <- pindex %>%
  filter(session == 2014 & time == 7 & Plot %in% as.numeric(names(plotssurveyed[,7,2])[which(plotssurveyed[,7,2] == 0)])) 
p2014.ind8 <- pindex %>%
  filter(session == 2014 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,2])[which(plotssurveyed[,8,2] == 0)])) 
p2014.ind9 <- pindex %>%
  filter(session == 2014 & time == 9 & Plot %in% as.numeric(names(plotssurveyed[,9,2])[which(plotssurveyed[,9,2] == 0)])) 
p2014.ind10 <- pindex %>%
  filter(session == 2014 & time == 10 & Plot %in% as.numeric(names(plotssurveyed[,10,2])[which(plotssurveyed[,10,2] == 0)])) 
p2014.inds <- c(p2014.indices$par.index, p2014.ind1$par.index, p2014.ind2$par.index, p2014.ind3$par.index, p2014.ind5$par.index,
                p2014.ind6$par.index, p2014.ind7$par.index, p2014.ind8$par.index, p2014.ind9$par.index, p2014.ind10$par.index)

## Missed in 2015
p2015.indices <- pindex %>%
  filter(session == 2015 & time == 10) 
p2015.ind1 <- pindex %>%
  filter(session == 2015 & time == 1 & Plot %in% as.numeric(names(plotssurveyed[,1,3])[which(plotssurveyed[,1,3] == 0)])) 
p2015.ind2 <- pindex %>%
  filter(session == 2015 & time == 2 & Plot %in% as.numeric(names(plotssurveyed[,2,3])[which(plotssurveyed[,2,3] == 0)])) 
p2015.ind3 <- pindex %>%
  filter(session == 2015 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,3])[which(plotssurveyed[,3,3] == 0)]))
p2015.ind4 <- pindex %>%
  filter(session == 2015 & time == 4 & Plot %in% as.numeric(names(plotssurveyed[,4,3])[which(plotssurveyed[,4,3] == 0)]))
p2015.ind5 <- pindex %>%
  filter(session == 2015 & time == 5 & Plot %in% as.numeric(names(plotssurveyed[,5,3])[which(plotssurveyed[,5,3] == 0)])) 
p2015.ind6 <- pindex %>%
  filter(session == 2015 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,3])[which(plotssurveyed[,6,3] == 0)])) 
p2015.ind7 <- pindex %>%
  filter(session == 2015 & time == 7 & Plot %in% as.numeric(names(plotssurveyed[,7,3])[which(plotssurveyed[,7,3] == 0)])) 
p2015.ind8 <- pindex %>%
  filter(session == 2015 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,3])[which(plotssurveyed[,8,3] == 0)])) 
p2015.ind9 <- pindex %>%
  filter(session == 2015 & time == 9 & Plot %in% as.numeric(names(plotssurveyed[,9,3])[which(plotssurveyed[,9,3] == 0)])) 

p2015.inds <- c(p2015.indices$par.index, p2015.ind1$par.index, p2015.ind2$par.index, p2015.ind3$par.index, p2015.ind4$par.index,
                p2015.ind5$par.index, p2015.ind6$par.index, p2015.ind7$par.index, p2015.ind8$par.index, p2015.ind9$par.index)


## Final for all missed ones that need to be zero
p.indices <- c(p2013.inds,p2014.inds, p2015.inds)
p.values <- rep(0, length(p.indices))

############################### c indeces
cindex <- RDddl$c
## Missed entire survey period
c2013.indices <- cindex %>%
  filter(session == 2013 & time %in% c(1,2,4,5,7,9))
## Missed these plots [Plot, Survey, Year]
c2013.ind3 <- cindex %>%
  filter(session == 2013 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,1])[which(plotssurveyed[,3,1] == 0)]))
c2013.ind6 <- cindex %>%
  filter(session == 2013 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,1])[which(plotssurveyed[,6,1] == 0)])) 
c2013.ind8 <- cindex %>%
  filter(session == 2013 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,1])[which(plotssurveyed[,8,1] == 0)])) 
c2013.ind10 <- cindex %>%
  filter(session == 2013 & time == 10 & Plot %in% as.numeric(names(plotssurveyed[,10,1])[which(plotssurveyed[,10,1] == 0)])) 
c2013.inds <- c(c2013.indices$par.index, c2013.ind3$par.index, c2013.ind6$par.index, c2013.ind8$par.index, c2013.ind10$par.index)

## Missed in 2014
c2014.indices <- cindex %>%
  filter(session == 2014 & time == 4)
c2014.ind1 <- cindex %>%
  filter(session == 2014 & time == 1 & Plot %in% as.numeric(names(plotssurveyed[,1,2])[which(plotssurveyed[,1,2] == 0)])) 
c2014.ind2 <- cindex %>%
  filter(session == 2014 & time == 2 & Plot %in% as.numeric(names(plotssurveyed[,2,2])[which(plotssurveyed[,2,2] == 0)])) 
c2014.ind3 <- cindex %>%
  filter(session == 2014 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,2])[which(plotssurveyed[,3,2] == 0)])) 
c2014.ind5 <- cindex %>%
  filter(session == 2014 & time == 5 & Plot %in% as.numeric(names(plotssurveyed[,5,2])[which(plotssurveyed[,5,2] == 0)]))
c2014.ind6 <- cindex %>%
  filter(session == 2014 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,2])[which(plotssurveyed[,6,2] == 0)])) 
c2014.ind7 <- cindex %>%
  filter(session == 2014 & time == 7 & Plot %in% as.numeric(names(plotssurveyed[,7,2])[which(plotssurveyed[,7,2] == 0)])) 
c2014.ind8 <- cindex %>%
  filter(session == 2014 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,2])[which(plotssurveyed[,8,2] == 0)])) 
c2014.ind9 <- cindex %>%
  filter(session == 2014 & time == 9 & Plot %in% as.numeric(names(plotssurveyed[,9,2])[which(plotssurveyed[,9,2] == 0)])) 
c2014.ind10 <- cindex %>%
  filter(session == 2014 & time == 10 & Plot %in% as.numeric(names(plotssurveyed[,10,2])[which(plotssurveyed[,10,2] == 0)])) 
c2014.inds <- c(c2014.indices$par.index, c2014.ind1$par.index, c2014.ind2$par.index, c2014.ind3$par.index, c2014.ind5$par.index,
                c2014.ind6$par.index, c2014.ind7$par.index, c2014.ind8$par.index, c2014.ind9$par.index, c2014.ind10$par.index)


## Missed in 2015
c2015.indices <- cindex %>%
  filter(session == 2015 & time == 10) 
c2015.ind1 <- cindex %>%
  filter(session == 2015 & time == 1 & Plot %in% as.numeric(names(plotssurveyed[,1,3])[which(plotssurveyed[,1,3] == 0)])) 
c2015.ind2 <- cindex %>%
  filter(session == 2015 & time == 2 & Plot %in% as.numeric(names(plotssurveyed[,2,3])[which(plotssurveyed[,2,3] == 0)])) 
c2015.ind3 <- cindex %>%
  filter(session == 2015 & time == 3 & Plot %in% as.numeric(names(plotssurveyed[,3,3])[which(plotssurveyed[,3,3] == 0)]))
c2015.ind4 <- cindex %>%
  filter(session == 2015 & time == 4 & Plot %in% as.numeric(names(plotssurveyed[,4,3])[which(plotssurveyed[,4,3] == 0)]))
c2015.ind5 <- cindex %>%
  filter(session == 2015 & time == 5 & Plot %in% as.numeric(names(plotssurveyed[,5,3])[which(plotssurveyed[,5,3] == 0)])) 
c2015.ind6 <- cindex %>%
  filter(session == 2015 & time == 6 & Plot %in% as.numeric(names(plotssurveyed[,6,3])[which(plotssurveyed[,6,3] == 0)])) 
c2015.ind7 <- cindex %>%
  filter(session == 2015 & time == 7 & Plot %in% as.numeric(names(plotssurveyed[,7,3])[which(plotssurveyed[,7,3] == 0)])) 
c2015.ind8 <- cindex %>%
  filter(session == 2015 & time == 8 & Plot %in% as.numeric(names(plotssurveyed[,8,3])[which(plotssurveyed[,8,3] == 0)])) 
c2015.ind9 <- cindex %>%
  filter(session == 2015 & time == 9 & Plot %in% as.numeric(names(plotssurveyed[,9,3])[which(plotssurveyed[,9,3] == 0)])) 

c2015.inds <- c(c2015.indices$par.index, c2015.ind1$par.index, c2015.ind2$par.index, c2015.ind3$par.index, c2015.ind4$par.index,
                c2015.ind5$par.index, c2015.ind6$par.index, c2015.ind7$par.index, c2015.ind8$par.index, c2015.ind9$par.index)
## Final for all missed ones that need to be zero
c.indices <- c(c2013.inds,c2014.inds, c2015.inds)
c.values <- rep(0, length(c.indices))

Huggins <- mark(data = RDproc, ddl = RDddl, model = "CRDMSOHug",
                      time.intervals = time.intervals, begin.time = 2013)

Huggins$results$real %>%
  distinct(estimate, se, lcl, ucl)

HugginsRDclosedMSmodeldot <- mark(data = RDproc, ddl = RDddl, model = "CRDMSOHug",
                      time.intervals = time.intervals,
                      model.parameters = list(
                                              # S = list(formula = ~ stratum + Plot),
                                              # Psi = list(formula = ~ stratum + Plot),
                                              # Omega = list(formula = ~ stratum + Plot),
                                              S = list(formula = ~ stratum ),
                                              Psi = list(formula = ~ stratum ),
                                              Omega = list(formula = ~ stratum ),
                                              p = list(formula = ~ stratum, fixed = list(index = p.indices, value = p.values)),
                                              c = list(formula = ~ stratum, fixed = list(index = c.indices, value = c.values)))
                      )

head(summary(HugginsRDclosedMSmodeldot, show.fixed = TRUE))
HugginsRDclosedMSmodeldot$results$real %>%
  distinct(estimate, se, lcl, ucl)

plotSite <- obsmodel %>%
  distinct(Plot, Site) %>%
  dplyr::rename(plot = Plot) %>%
  mutate(plot = as.character(plot))
```


